---
title: "pca test"
author: "Eduardo Flores"
date: "9/14/2017"
output: html_document
---

# Question: When transforming vectors to their Principal Components, are their relations preserved?

We have a set of vectors. After performing [Principal Component Analysis (PCA)](https://en.wikipedia.org/wiki/Principal_component_analysis) we now use the "rotated" vectors to perform analysis. **Can we be confident that the original relations (cosine similarity between vectors) are preserved on the new vector space?**

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1234)
#Cosine Similarity
cos.sim <- function(A,B) 
{
  return( sum(A*B)/sqrt(sum(A^2)*sum(B^2)) )
}   
```
### 1 - Generate a test set y=x+err

```{r}
#"noisy" x=y
x = runif(n = 1000,min=-100,max=100) # x has a uniform distribution
y = x + rnorm(n=1000,mean=0,sd=20) # y has a normal distribution

d <- data.frame(x=x,y=y)
#ver primeros puntos
head(d)
plot(d)
```

### 2 - Perform PCA
```{r}
pca <- prcomp(d,center = F,scale. = F)
summary(pca)
pca$rotation
dt <- pca$x #d transformed
head(dt)
plot(dt)
```

### 3 - For a given vector, calculate it's relation (cosine similarity) to all the others in both spaces
```{r}
a <- d[2,] #choose a point on the original space
s<-apply(d,1,function(x) cos.sim(a,x)) #calculate the angles to all the other vectors
at <- dt[2,] #get the same point in the new space
st<-apply(dt,1,function(x) cos.sim(at,x)) #calculate the angles in the new space
```
#### The angles are the same
```{r}
head(data.frame(s=s,st=st))
```
#### Check closest points are the same
```{r}
head(data.frame(s=order(s,decreasing = T),st=order(st,decreasing = T)))
```

# Answer: YES

### 4 - Bonus: magnitudes are also the same

```{r}
magd <- sqrt(rowSums(d*d))
magdt <- sqrt(rowSums(dt*dt))

head(data.frame(magd=magd,magdt=magdt))
```

### 5 - Bonus: you can retreive the distributions 

```{r}
hist(as.data.frame(dt)$PC1) #this is close to uniform
hist(as.data.frame(dt)$PC2) #this is close to normal
```

### Looks closer than the original x,y

```{r}
hist(d$x)
hist(d$y)
```


